---
title: "Lab 7 Final"
author: "Harshini Karthikeyan"
date: "December 3, 2022"
mainfont: "Times New Roman"
format: 
  html:
    code-fold: true
    code-summary: "Show code"
    monobackgroundcolor: lightsteelblue
    backgroundcolor: whitesmoke
    fontcolor: steelblue
    toc: true
    theme: zephyr
    self-contained: true
execute: 
  warning: false
  message: false
editor: visual
---

### Data

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
library(patchwork)
blackfoot_fish <- read_csv(here::here("supporting_artifacts",
                                    "Labs", "Lab_7_Final",
                                    "BlackfootFish.csv"))

```

### Part One: Summaries and Plots

Task 1 - Summarizing Missing Data

```{r summarizing missing data}

blackfoot_fish |> 
  summarize(across(.cols = everything(), ~sum(is.na(.))))
```

1796 observations have missing values and weight is the variable with missing values.

Task 2 - Visualizing Missing Data

```{r visualizing missing data}
blackfoot <- blackfoot_fish |> 
   mutate("num" = across(.cols = weight,
                 .fns = ~(is.na(.)))) |> 
  group_by(year, species, section)|>
  summarize(across(.cols = everything(), ~sum(.)), .groups = "rowwise")

ggplot(blackfoot) +
  geom_point(mapping = aes(x = year,
                y = num,
                color = fct_rev(reorder(species,
                                        num)))) +
  facet_wrap(~section) +
  theme( plot.title.position = "plot",
         plot.title = element_text(vjust = 1,
                                   family = "serif")) +
  labs( title = "Number of Missing Values in Species Weight over Time
        by Section",
        x = " ", y = " ", color = "Species Name")
```

### Part Two: Adjusting the Data (Function Writing)

Task 1 & 2 - Writing a Function + Adding Stops

```{r tasks_1_and_2}
rescale_01 <- 
  function(vec){
    stopifnot(is.numeric(vec),
              length(vec) > 1)
    mins <- min(vec, na.rm = TRUE)
    maxes <- max(vec, na.rm = TRUE)
    vec2 <- ((vec - mins) / (maxes - mins))
    return(vec2)
  }

```

\
Task 3 - Performing a Simple Test

```{r task_3}
x <- c(1:25, NA)
x1 <- rescale_01(x) 

max(x1, na.rm = TRUE)

min(x1, na.rm = TRUE)
```

Task 4 - Performing a More Difficult Test

```{r task_4}
blackfoot_length <- blackfoot_fish |>
  mutate('rescaled_length' = rescale_01(vec = length)) |>
  group_by(year) |> summarise(length = mean(length),
                              rlen = mean(rescaled_length))


len <- 
  ggplot(data = blackfoot_length,
         mapping = aes(x = year,
                       y = length)) +
  geom_line(color = 'chartreuse4') +  
   theme(plot.title.position = "plot",
         plot.title = element_text(family = "serif")) +
  labs( title = "Length of Blackfoot Fish across Time
        ",
        x = " ", y = " ")

len1 <- 
  ggplot(data = blackfoot_length,
         mapping = aes(x = year,
                       y = rlen)) +
  geom_line(color = 'cyan4')  +  
   theme(plot.title.position = "plot",
         plot.title = element_text(family = "serif")) +
  labs( title = "Rescaled Length of Blackfoot Fish across
        Time",
        x = " ", y = " ")

len + len1
```

Task 5 -- Incorporating Variables

```{r task_5}
rescale_column <- 
  function(data_name, col_names){
    data2 <- data_name |> 
      mutate(across(.cols = {{col_names}},
             .fns =  ~rescale_01(.)))
    return(data2)
  }
```

Task 6 -- Another Function Test

```{r task_6}
#| message: false
rescale_column(blackfoot_fish, c(length, weight))

```

REVISIONS:

In Task 1 of Section One I was performing an unnecessary side step by creating a new column before summarizing to produce the summary graph when I could have directly summarized the data.

In Task 2 of Section One I connected the data points in the graph using geom_line() which was not the most accurate portrayal of the data, therefore I switched it to geom_point().

In Task 5 and 6 of Section Two, I forgot to add the two {{}} brackets around the vector containing the column names, in the function, so that my column names would not need to be in quotations.
